---
title: "assignment-b4"
author: "Alma Garcia"
date: "2023-12-04"
output: github_document
---

# Strings and functional programming in R

## Exercise 1

For this exercise I chose the book "Frankenstein" by Mary Shelley (a favorite of mine). The book was downloaded as plain text from [Project Gutenberg](https://www.gutenberg.org/ebooks/84). 

```{r}
library(stringr)
library(purrr)
library(dplyr)
library(tm)
```

```{r}
con <- file("files/frankenstein.txt")
frankenstein_book <- readLines(con) %>% tolower()
head(frankenstein_book)
```

So the book actually starts at line 70 and ends at line 7385, we will extract the relevant content to work with it. I will also keep lines that have at least 1 character. 

```{r}
frankenstein <- frankenstein_book[70:7385] %>%
	str_subset(".+")

head(frankenstein)
```
Now we need to create our unique words list. 

```{r}
frankenstein_words <- frankenstein %>% 
	str_split("\\s+") %>% # split lines into words
	unlist() %>% # break the list structure
	str_replace_all("[:punct:]", " ") %>% # turns punctuation into spaces (because some words have em dashes between)
	str_subset("^[a-z]+$") # filters out punctuation and numbers
```

```{r}
head(frankenstein_words, 20)
```
And then we need to count the number of unique words so we can create plots:

```{r}
frankenstein_unique <- as_tibble(frankenstein_words) %>%
	filter(!value %in% tm::stopwords("SMART")) %>% # remove stop words
	group_by(value) %>%
	reframe(count = n()) %>%
	rename(word = value) 

head(frankenstein_unique)
```

To visualize the most used words let's make a word cloud using the `wordcloud` package:

```{r}
library(wordcloud)
```

```{r}
# color palette
pal = c("#3C4B69FF", "#4B5A69FF", "#697878FF", "#4B4B4BFF", "#3C3C3CFF", "#3C3C1EFF", "#695A2DFF", "#78693CFF", "#78784BFF", "#C0BB9DFF", "#C2902FFF", "#965A0FFF", "#784B0FFF", "#5A4B0FFF")
```


```{r}
wordcloud(frankenstein_unique$word, 
					frankenstein_unique$count, 
					max.words = 500,
					random.order = FALSE, 
					random.color = TRUE,
					scale = c(2.5, 0.05), 
					rot.per=0,
					fixed.asp = FALSE,
					family = "Kingthings Trypewriter 2", 
					font = 1, 
					col = pal)
```

And we'll make another plot, this time a barplot, to show the top 15 words:

```{r}
frankenstein_unique %>%
	arrange(desc(count)) %>%
	slice_head(n = 15) %>%
	ggplot(aes(reorder(word, -count), count)) +
	geom_col(fill = pal[1], position = position_dodge(width=0.5)) +
	xlab("word") +
	scale_y_continuous(expand = c(0,0.15)) +
	theme_minimal() +
	theme(text = element_text(color = pal[1], family = "Kingthings Trypewriter 2"),
				axis.title = element_text(size = 16),
				axis.text = element_text(color = pal[1]),
				axis.text.x = element_text(angle = 30, hjust=1),
				axis.ticks.x.bottom = element_line(linewidth = 0.5, color = "#EBEBEB"),
				panel.grid.major.x = element_blank()
				) 
```
